---
title: "Statistical Inference: Exponential Distribution"
author: Ariel Lev
date: "26. April 2015"
output: 
  html_document:
    keep_md: true
---
#### Overview
In this report I would like to emperically show the properties of the Central Limit Theorem given samples of exponential distrubted iid variables. 
I will show that the estimates of the mean and variance converge respectively to the population's mean and variance, and that the distribution of the sample converge to a normal distribution. 

#### Simulations
```{r,echo=T, cache=T}
# number of exponentials in each simulation
n <- 40
# number of simulations
nsim <- 1000
# lambda
lambda <- 0.2

# mean and variance vectors
mns = NULL
vars = NULL

# theoretical mean and variance
data <- rexp(n*nsim, lambda)
data.frame( mean = mean(data), var = var(data))

## simulation begins
for (i in 1:nsim) {
  r_expos <- rexp(n, lambda)
  mns <- c(mns,mean(r_expos))
  vars <- c(vars, var(r_expos))
  }
```


#### Sample Mean vs. Theoretical Mean
```{r,echo=F, cache=T, message=F, warning=F}
require(ggplot2)
```
```{r,echo=T, cache=T, message=F, warning=F, fig.width=12, fig.height=6}
expo_df <-data.frame( value = data, type = "Theoretical")
mns_df <-data.frame( value = mns, type = "Sample")
plot_df <- rbind(expo_df, mns_df)

data.frame( mean = c(mean(data), mean(mns)), row.names = c("theoretical", "sample"))

ggplot(as.data.frame(plot_df$value),aes(x=plot_df$value)) + geom_density(aes(fill=plot_df$type, alpha = 1)) + geom_vline(xintercept = mean(expo_df$value), color = "red", size = 1) + geom_vline(xintercept = mean(mns_df$value), linetype = "longdash", color = "blue", size = 0.5) + scale_x_continuous(breaks = sort(seq(0,20,5))) + ggtitle(label = "Sample Mean vs. Theoretical Mean") + theme(axis.title.x=element_blank()) + guides(fill=guide_legend(title="Type")) + xlim(0,20)
```

#### Sample Variance vs. Theoretical Variance
```{r,echo=T, cache=T, message=F, fig.width=12, fig.height=6}
expo_df <-data.frame( value = data, type = "Theoretical")
mns_df <-data.frame( value = mns, type = "Sample")
plot_df <- rbind(expo_df, mns_df)

# theoretical variance and mean of sample variance
data.frame( var = c(var(data), mean(vars)), row.names = c("theoretical", "sample"))

# theoretical variance of samples 
var(data)/(n-1)

# empirical variance of samples 
var(mns)

```
#### Distribution
```{r,echo=T, cache=T, message=F, fig.width=12, fig.height=6}

normal_df <-data.frame( value = rnorm(nsim, mean = mean(mns), sd = sd(mns)), type = "Normal")
mns_df <-data.frame( value = mns, type = "Sample")
plot_df <- rbind(normal_df, mns_df)

ggplot(as.data.frame(plot_df$value),aes(x=plot_df$value)) + geom_density(aes(fill=plot_df$type, alpha = 1)) + geom_vline(xintercept = mean(normal_df$value), color = "red", size = 1) + geom_vline(xintercept = mean(mns_df$value), linetype = "longdash", color = "blue", size = 0.5) + scale_x_continuous(breaks = sort(seq(0,20,5))) + ggtitle(label = "Sample vs. Normal Distribution") + theme(axis.title.x=element_blank()) + guides(fill=guide_legend(title="Type")) + xlim(0,20)
```